### 2021-07-15 chapters/1_abstract.tex
chagnes

\documentclass[../main]{subfiles}
\begin{document}
\chapter*{Abstract}
Software testing is a crucial part of the software engineering process. 
It is widely used in the industry.
A part of software testing is building test suites which contain unit tests.
These unit tests are written by developers. 
As projects grow the test suite grows along.
Maintaining and monitoring these test suites is important because they influence the cost of maintenance.
Monitoring the quality of test code has shown that tests with lower quality leads to more defect-prone production code.
\newline
Measuring the effectiveness of test suites detecting faults is part of the monitoring process.
A technique to measure the test effectiveness is mutation testing.
Mutation testing is a computationally expensive technique that requires the ability to run tests.
Different techniques have been proposed to reduce the cost of mutation testing.
One of these techniques is the clustering of mutants.
By clustering mutants we can execute less mutants to reduce the cost.
\newline
We conducted two experiment in which we cluster the mutants with two different techniques.
Before conducting the experiment we conducted a state of the art comparison between three mutation testing tools for Java.
Based on this research we selected a mutation testing tool to use in our experiments.
\newline
We started our experiment by identifying characteristics to represent the mutants.
For the first experiment we used hierarchical clustering and in the second experiment we trained the \acrlong{fcm} model.
We calculated a weighted mutation score and compared this with the mutation score of a full set of mutants executed.
Results show that with hierarchical clustering we can reduce the amount of mutants executed while maintaining the effectiveness.
The clusters generated by the machine learning model were less accurate and showed a significant decrease in effectiveness.
The machine learning model didn't perform as well as the hierarchical clustering.
However, more work is needed in finding the right parameters for the \acrlong{fcm} algorithm. 
\end{document}

>>>>

\documentclass[../main]{subfiles}
\begin{document}
\chapter*{Abstract}
Software testing is a crucial part of the software engineering process. 
It is widely used in the industry.
A part of software testing is building test suites which contain unit tests.
These unit tests are written by developers. 
As projects grow the test suite grows along.
Maintaining and monitoring these test suites is important because they influence the cost of maintenance.
Monitoring the quality of test code has shown that tests with lower quality lead to more defect-prone production code.
\newline
Measuring the effectiveness of test suites detecting faults is part of the monitoring process.
A technique to measure the test effectiveness is mutation testing.
Mutation testing is a computationally expensive technique that requires the ability to run tests.
Different techniques have been proposed to reduce the cost of mutation testing.
One of these techniques is the clustering of mutants.
By clustering mutants we can execute less mutants to reduce the cost.
\newline
We conducted two experiments in which we cluster the mutants with two different techniques.
Before conducting the experiment we conducted a state of the art comparison between three mutation testing tools for Java.
Based on this research we selected a mutation testing tool to use in our experiments.
\newline
We started our experiment by identifying characteristics to represent the mutants.
For the first experiment we used hierarchical clustering and in the second experiment we trained the \acrlong{fcm} model.
We calculated a weighted mutation score and compared this with the mutation score of a full set of mutants executed.
Results show that with hierarchical clustering we can reduce the amount of mutants executed while maintaining the effectiveness.
The clusters generated by the machine learning model were less accurate and showed a significant decrease in effectiveness.
The machine learning model did not perform as well as the hierarchical clustering.
However, more work is needed in finding the right parameters for the \acrlong{fcm} algorithm. 
\end{document}

### 2021-07-15 chapters/2_introduction.tex
chagnes

\chapter{Introduction}
\label{ch:introduction}
Software testing is a critical part in the software engineering process. It is widely used and accepted in the industry for quality assurance. 
Software testing can detect software bugs during the development process and can also serve for regression purposes\cite{Bertolino2007SoftwareDreams}. 
During the software testing process developers write automated tests called unit tests. 
These tests can be run by testing frameworks such as JUnit and TestNG\cite{JUnit, TestNG}.
Monitoring the quality of test code has shown that tests with lower quality leads to more defect-prone production code\cite{Spadini2018OnQuality}. 
Research has also found that if the production code grows, the test code grows along with it\cite{Zaidman2008MiningCode}. 
Quality control on test suites is therefore important as the maintenance on tests can be difficult and generates risks if done incorrectly\cite{Horvath2015TestMetrics}.
\newline
The growing complexity and size of code bases consequently lead to incomprehensible tests.
As a result test bugs can occur. 
A test bug is a test that fail even tough the program is correct(\textit{false positive}). 
Another bug, which is even worse, are tests that do not fail when the program is not working as desired(\textit{false negative}).
The \textit{false negative} is a problem when breaking changes are not detected by the test suite.
\newline
To solve this issue we can measure the fault detecting capability of a test suite, this is called the test suite effectiveness.
Test suite effectiveness is measured by the number of faulty versions of a \acrfull{sut} that are detected by a test suite.
Real faults are unknown in advance, we can use mutation testing as a proxy measurement. 
Research has found statistically significant evidence which correlates mutant detection with real fault detection\cite{Just2014AreTesting}.
\newline
Mutation testing tools create small faulty versions of the program.
The test suite is then run to verify if these faults are detected.
These faults are called mutants and are created by mutators.
Mutators mutate specific expressions or statements in the source code. 
To prevent changing the overall functionality of the program, mutants are represented as very small changes\cite{Lipton1978HintsProgrammer}.
\newline
After all the mutants are executed the mutation score can be calculated. 
The mutation score is the percentage of mutants killed divided by the total amount of mutants executed.
If a large amount of mutants survives, it might be an indication that the quality of the test suite is poor as programming errors remain undetected.
Mutation testing has considerable drawbacks, such as the equivalent mutant problem and being resource expensive.

\section{Problem statement}
Mutation testing requires a significant amount of computational resources\cite{Kintis2016EffectiveMutation}. 
The high cost requirement is often a barrier for adopting mutation testing \cite{Pizzoleto2019}.
A lot of techniques and methods have been developed to improve the performance, however, most of these approaches are not as effective as mutation testing a full set of mutants\cite{Pizzoleto2019,Yao2014}. 
\newline
A technique to reduce the number of mutants or the number of test case executions is mutation clustering.
Mutation clustering aims to reduce the amount of mutants to be executed by clustering the mutants\cite{Ma2016,Yu2019PossibilityScope}.
Clustering is a technique used for the grouping of data, for pattern recognition and classification, based on some similar characteristics\cite{Alsabti1997}.
The clustering of mutants has been researched with promising results\cite{Ji2009,Wilinski2015,Ma2016}. For example Ma et al., \cite{Ma2016} clustered the mutants for expressions. 
However in their paper they indicate that the choosing expressions as the domain of their cluster is a limitation.
A recurring problem in the research of mutation clustering is to determine the means of domains as the centre of a cluster\cite{Ji2009,Wilinski2015,Ma2016,Wei2021SpectralTesting}.
\newline
To reduce the cost of mutation testing, we try to find a solution that can cluster mutants while maintaining the accuracy of the same complete set of mutants. 
Our goal is to cluster any mutant that is generated.
We will do this with two different approaches.
Our first approach will be a qualitative approach and second a quantitative approach. 
We structure our research on the following research questions:
\newline
\textbf{Research Question 1}: What set of characteristics can we identify for clustering generated mutants to reduce the amount executed, while maintaining effectiveness?
\newline
\textbf{Research Question 1.1}: How do the existing mutation testing tools for Java compare to each other?
\newline
\textbf{Research Question 2}: How can we train a machine learning model to recognize and cluster generated mutants to reduce the amount executed, while maintaining effectiveness?
\newline
We start with comparing existing mutation testing tools. 
We will use the existing literature as a starting point and research what would be the best tool to use for this thesis.
\newline
After we have selected a tool we will research existing clustering techniques.
Based on the results of this research we select and implement a clustering algorithm that will cluster one ore more characteristics of the mutants. 
We will randomly select on mutant per of each cluster to be executed by the mutation tooling.
We will then compare the mutation score of the clustered mutants with the mutation score of a full set of mutants.
\newline
To answer our second research question we will train a machine learning model to cluster mutants.
We can then repeat the comparison we did for research question one. 
\newline
We want to see how the algorithm/model performs on big and small projects.
These project must also require a test suite.
\newline
We chose to focus on projects that use Java as main programming language because it is one of the most popular languages\cite{Tiobe}.

\section{Contributions}
Our research makes the following contributions:
\begin{enumerate}
 \item A methodology to cluster generated mutants based on chosen characteristics to reduce the cost of mutation testing.
 \item A \acrfull{poc} which implements the qualitative methodology chosen and elaborated within the thesis.
 \item A \acrfull{poc} which implements the quantitative approach to cluster mutants and is elaborated within the thesis. 
\end{enumerate}

>>>>

\chapter{Introduction}
\label{ch:introduction}
Software testing is a critical part in the software engineering process. It is widely used and accepted in the industry for quality assurance. 
Software testing can detect software bugs during the development process and can also serve for regression purposes\cite{Bertolino2007SoftwareDreams}. 
During the software testing process developers write automated tests called unit tests. 
These tests can be run by testing frameworks such as JUnit and TestNG\cite{JUnit, TestNG}.
Monitoring the quality of test code has shown that tests with lower quality leads to more defect-prone production code\cite{Spadini2018OnQuality}. 
Research has also found that if the production code grows, the test code grows along with it\cite{Zaidman2008MiningCode}. 
Quality control on test suites is therefore important as the maintenance on tests can be difficult and generates risks if done incorrectly\cite{Horvath2015TestMetrics}.
\newline
The growing complexity and size of code bases consequently lead to incomprehensible tests.
As a result test bugs can occur. 
A test bug is a test that fail even tough the program is correct(\textit{false positive}). 
Another bug, which is even worse, are tests that do not fail when the program is not working as desired(\textit{false negative}).
The \textit{false negative} is a problem when breaking changes are not detected by the test suite.
\newline
To solve this issue we can measure the fault detecting capability of a test suite, this is called the test suite effectiveness.
Test suite effectiveness is measured by the number of faulty versions of a \acrfull{sut} that are detected by a test suite.
Real faults are unknown in advance, we can use mutation testing as a proxy measurement. 
Research has found statistically significant evidence which correlates mutant detection with real fault detection\cite{Just2014AreTesting}.
\newline
\todo[inline]{is this your approach? the current general approach?}
Mutation testing tools create small faulty versions of the program.
The test suite is then run to verify if these faults are detected.
These faults are called mutants and are created by mutators.
Mutators mutate specific expressions or statements in the source code. 
To prevent changing the overall functionality of the program, mutants are represented as very small changes\cite{Lipton1978HintsProgrammer}.
\newline
After all the mutants are executed the mutation score can be calculated. 
The mutation score is the percentage of mutants killed divided by the total amount of mutants executed.
If a large amount of mutants survives, it might be an indication that the quality of the test suite is poor as programming errors remain undetected.
Mutation testing has considerable drawbacks\todo{add ref}, such as the equivalent mutant problem and being resource expensive.

\section{Problem statement}
Mutation testing requires a significant amount of computational resources\cite{Kintis2016EffectiveMutation}. 
The high cost requirement is often a barrier for adopting mutation testing \cite{Pizzoleto2019}.
A lot of techniques and methods have been developed to improve the performance, however, most of these approaches are not as effective as mutation testing a full set of mutants\cite{Pizzoleto2019,Yao2014}. 
\newline
A technique to reduce the number of mutants or the number of test case executions is mutation clustering.
Mutation clustering aims to reduce the amount of mutants to be executed by clustering the mutants\cite{Ma2016,Yu2019PossibilityScope}.
Clustering is a technique used for the grouping of data, for pattern recognition and classification, based on some similar characteristics\cite{Alsabti1997}.
The clustering of mutants has been researched with promising results\cite{Ji2009,Wilinski2015,Ma2016}. For example Ma et al., \cite{Ma2016} clustered the mutants for expressions. 
However in their paper they indicate that the choosing expressions as the domain of their cluster is a limitation.
A recurring problem in the research of mutation clustering is to determine the means of domains as the centre of a cluster\cite{Ji2009,Wilinski2015,Ma2016,Wei2021SpectralTesting}.
\newline
To reduce the cost of mutation testing, we try to find a solution that can cluster mutants while maintaining the accuracy of the same complete set of mutants. 
Our goal is to cluster any mutant that is generated \todo{make the goal clearer: no requirements on which mutants get clustered: aim at generic mutant clustering}.
We will do this with two different approaches: qualitative and quantitative. 
We structure our research as follows:
\newline
\textbf{Research Question 1}: What set of characteristics can we identify for clustering generated mutants to reduce the amount executed, while maintaining effectiveness?
\newline
\textbf{Research Question 1.1}: How do the existing mutation testing tools for Java compare to each other?
\newline
\textbf{Research Question 2}: How can we train a machine learning model to recognize and cluster generated mutants to reduce the amount executed, while maintaining effectiveness?
\newline
We start with comparing existing mutation testing tools. 
We will use the existing literature as a starting point and research what would be the best tool to use for this thesis.
\newline
After we have selected a tool we will research existing clustering techniques.
Based on the results of this research we select and implement a clustering algorithm that will cluster one ore more characteristics of the mutants. 
We will randomly select on mutant per of each cluster to be executed by the mutation tooling.
We will then compare the mutation score of the clustered mutants with the mutation score of a full set of mutants.
\newline
To answer our second research question we will train a machine learning model to cluster mutants.
We can then repeat the comparison we did for research question one. 
\newline
We want to see how the algorithm/model performs on big and small projects.
These project must also require a test suite.
\newline
We chose to focus on projects that use Java as main programming language because it is one of the most popular languages\cite{Tiobe}.

\section{Contributions}
Our research makes the following contributions:
\begin{enumerate}
 \item A methodology to cluster generated mutants based on chosen characteristics to reduce the cost of mutation testing.\todo{both white box and black box?}
 \item A \acrfull{poc} which implements the qualitative methodology chosen and elaborated within the thesis.
 \item A \acrfull{poc} which implements the quantitative approach to cluster mutants and is elaborated within the thesis. 
\end{enumerate}

### 2021-07-15 chapters/rq2/1_experiment_design.tex
chagnes

\subsection{Implementation of algorithm}
The research paper of the \acrshort{fcm} algorithm is referenced in the survey of Du et al\cite{Du2010Clustering:Approach}.
The \acrshort{fcm} algorithm is developed by Bezdek et al.\cite{Bezdek1984FCM:Algorithm}.
Based on their research an implementation can be developed.
There are existing libraries that contain the implementation of the \acrshort{fcm} algorithm.
Developing our own version of the \acrshort{fcm} algorithm would be valuable if we want to make adjustments to the algorithm itself.
For this experiment we do not require to make adjustments to the \acrshort{fcm} algorithm.
Therefore we choose to use a library that implements the \acrshort{fcm} algorithm.
The library we will use for this experiment is the \acrlong{fcm} implementation build by Dias et al.\cite{dias2019fuzzy}.
The theory developed by Bezdek et al., is implemented in this library.
The library is actively maintained and has been used before in other research\cite{DeAlmeidaNeto2020, Kopf2019, Nwadiugwu2020}.

\subsection{Algorithm parameters}
The \acrshort{fcm} algorithm has to initialize the centroids of the clusters.
The library we selected does this based on a random number.
It also allows us to control this random number.
To make the initialization of the centroid reproducible we will use the same generated seeds as in our first experiment.
The seeds generated are included in the Github repository\cite{rbasarat-repo}.
\newline
Bezdek et al., also researched cluster validity for the \acrshort{fcm} model\cite{Bezdek1995OnModel}.
They researched what effect which parameters have on the validity of the clusters.
They specifically analyzed the role of weighting exponent \textit{m}(fuzziness parameter).
They concluded that the best choice for \textit{m} is probably in the interval [1.5, 2.5], whose mean and midpoint is \textit{m}=2.
Wu proposed a new guideline for selecting parameter \textit{m}\cite{Wu2012}.
His point of view was that a large \textit{m} value will make the \acrlong{fcm} more robust to noise and outliers.
He suggests implementing \acrlong{fcm} with \textit{m}=[1.5, 4].
When the data set contains noise and outliers, the fuzzifier \textit{m}=4 is recommended in a theoretical upper bound case.
Since our data set does not contain noise and outliers we choose to select the midpoint concluded in the research of Bezdek et al.\cite{Bezdek1995OnModel}.
The value of \textit{m} in this experiment is 2.
\newline
Finding the correct amount of clusters has been an ongoing problem for clustering algorithms\cite{Du2010Clustering:Approach}.
Fortunately there are some techniques available to estimate the number of clusters.
During model selection we discussed the mountain and subtractive models. 
These models are designed to estimate the number of clusters in a data set.
The subtractive clustering model is specifically designed to approximate the number of clusters for fuzzy system models\cite{Chiu1994FuzzyEstimation}.
It approximates the centroids and amount of the clusters based on the data points in the data set\cite{Chiu1994FuzzyEstimation}.
We will select and implement the subtractive clustering model to estimate the number of clusters.
We can use this number for the number of clusters parameter in the \acrshort{fcm} algorithm.

\subsection{Implementation of subtractive clustering}
The research paper of the mountain and subtractive algorithms are referenced in the survey paper of Du et al\cite{Du2010Clustering:Approach}.
Based on the paper of the theory of the subtractive algorithm\cite{Chiu1994FuzzyEstimation} we found an existing implementation\cite{matlabSubtractive}.
The implementation is professionally build and maintained by Matlab\cite{matlabSubtractive}.
Matlab is a tool commonly used in the academic world and is professionally maintained for this reason we decided to use their implementation of subtractive clustering.

\subsection{Validation}
To validate our machine learning model we will calculate the weighted mutation score and cluster accuracy based on the clusters calculated by the model.
This is the same validation method as described in Chapter \ref{ch:exp1_validation}.
We will use the mutants of projects that are not selected for training as input for validating the model.
As described in Chapter \ref{ch:training_strategy} the projects should represent approximately 20\% of the total amount of mutants in our data set.
It may be possible the model has to cluster mutants it has never seen before.
We will use the same statistical significance threshold as in our firsts experiment to decide whether we will accept or reject or hypothesis.
To make our results reproducible we will select a random mutant based on a generated seed.
We will repeat our experiment 30 times with 30 different seeds.
Achieving consistent results while applying random selection and cluster initialization will contribute to the validity of the experiment.
The seed generated for each run will be included in the results and can be found in the Github repository\cite{rbasarat-repo}.

>>>>

\subsection{Implementation of algorithm}
\label{ch:fcm_implementation}
The research paper of the \acrshort{fcm} algorithm is referenced in the survey of Du et al\cite{Du2010Clustering:Approach}.
The \acrshort{fcm} algorithm is developed by Bezdek et al.\cite{Bezdek1984FCM:Algorithm}.
Based on their research an implementation can be developed.
There are existing libraries that contain the implementation of the \acrshort{fcm} algorithm.
Developing our own version of the \acrshort{fcm} algorithm would be valuable if we want to make adjustments to the algorithm itself.
For this experiment we do not require to make adjustments to the \acrshort{fcm} algorithm.
Therefore we choose to use a library that implements the \acrshort{fcm} algorithm.
The library we will use for this experiment is the \acrlong{fcm} implementation build by Dias et al.\cite{dias2019fuzzy}.
The theory developed by Bezdek et al., is implemented in this library.
The library is actively maintained and has been used before in other research\cite{DeAlmeidaNeto2020, Kopf2019, Nwadiugwu2020}.

\subsection{Algorithm parameters}
The \acrshort{fcm} algorithm has to initialize the centroids of the clusters.
The library we selected does this based on a random number.
It also allows us to control this random number.
To make the initialization of the centroid reproducible we will use the same generated seeds as in our first experiment.
The seeds generated are included in the Github repository\cite{rbasarat-repo}.
\newline
Bezdek et al., also researched cluster validity for the \acrshort{fcm} model\cite{Bezdek1995OnModel}.
They researched what effect which parameters could have on the validity of the clusters.
They specifically analyzed the role of weighting exponent \textit{m}(fuzziness parameter).
They concluded that the best choice for \textit{m} is in the interval [1.5, 2.5], whose mean and midpoint is \textit{m}=2.
Wu proposed a new guideline for selecting parameter \textit{m}\cite{Wu2012}.
His point of view was that a large \textit{m} value will make the \acrlong{fcm} more robust to noise and outliers.
He suggests implementing \acrlong{fcm} with \textit{m}=[1.5, 4].
When the data set contains noise and outliers, the fuzzifier \textit{m}=4 is recommended in a theoretical upper bound case.
Since our data set does not contain noise we choose to select the midpoint concluded in the research of Bezdek et al.\cite{Bezdek1995OnModel}.
The value of \textit{m} in this experiment is 2.
\newline
Finding the correct amount of clusters has been an ongoing problem for clustering algorithms\cite{Du2010Clustering:Approach}.
Fortunately there are techniques available to estimate the number of clusters.
During model selection we discussed the mountain and subtractive models. 
These models are designed to estimate the number of clusters in a data set.
The subtractive clustering model is specifically designed to approximate the number of clusters for fuzzy system models\cite{Chiu1994FuzzyEstimation}.
It approximates the centroids and amount of the clusters based on the data points in the data set\cite{Chiu1994FuzzyEstimation}.
We will select and implement the subtractive clustering model to estimate the number of clusters.
We will use this number as the value for the number of clusters parameter in the \acrshort{fcm} algorithm.

\subsection{Implementation of subtractive clustering}
For the same reasons as in Chapter \ref{ch:fcm_implementation} we will select a library that implements the subtractive clsutering algorithm.
The research paper of the mountain and subtractive algorithms are referenced in the survey paper of Du et al\cite{Du2010Clustering:Approach}.
Based on the paper of the theory of the subtractive algorithm\cite{Chiu1994FuzzyEstimation} we found an existing implementation\cite{matlabSubtractive}.
The implementation is professionally build and maintained by Matlab\cite{matlabSubtractive}.
Matlab is a tool commonly used in the academic world and is professionally maintained for this reason we decided to use their implementation of subtractive clustering.

\subsection{Validation}
To validate our machine learning model we will calculate the weighted mutation score and cluster accuracy based on the clusters calculated by the model.
We use the same validation method as described in Chapter \ref{ch:exp1_validation}.
We will use the mutants of projects that are not selected for training as input for validating the model.
As described in Chapter \ref{ch:training_strategy} the mutants should represent approximately 20\% of the total amount of mutants per project.
It may be possible the model has to cluster mutants it has never seen before.
We will use the same statistical significance threshold as in our firsts experiment to decide whether we will accept or reject or hypothesis.
To make our results reproducible we will select a random mutant based on a generated seed.
We will repeat our experiment 30 times with 30 different seeds.
Achieving consistent results while applying random selection and cluster initialization will contribute to the validity of the experiment.
The seed generated for each run will be included in the results and can be found in the Github repository\cite{rbasarat-repo}.
